Model Finding for Recursive Functions in SMT\thanks{ This research is partially supported by the Inria technological development action ``Contre-exemples Utilisables par Isabelle et Coq'' (CUIC). }

### 2015-05-07 tex/main.tex
more

Admissibility is a semantic criterion that must be satisfied for each function
definition before applying the translation described in
Section~\ref{sec:encoding}. This criterion is general enough to accommodate
\emph{well-founded} function definitions, also called ``terminating'' definitions in a
slight abuse of terminology. In such definitions, all recursive calls are
decreasing with respect to a well-founded relation $r$, which must be supplied
by the user or inferred automatically using a termination prover.

First-order logic has built-in notion of computation or termination. To ensure
that a function specification is well-founded, it is sufficient to require that
it would terminate when seen as a functional program for \relax{some}
evaluation order. For example,
$\{\forall n : {\typeint}.\;\, \lite\bigl(
n \leq 0,\allowbreak\; \con{p}( n ) \teq 1,\allowbreak\; \con{p}( n ) \teq 2 * \con{p}( n - 1 ) \bigr)\}$
can be shown well-founded under an evaluation strategy that consides one of the
two $\lite$ branches, based on the result of evaluating the condition.
Logically, such dependencies can be captured by congruence rules. Krauss
develops these ideas systematically in the more general context of higher-order
logic \cite[Section 2]{krauss-2009-phd}.

\begin{theorem}
Well-founded f
\end{theorem}

@@

The \texttt{--fmf-fun} option instructs \cvc to assume that the
functions provided by the user are admissible.
Admissibility must be checked separately \emph{by the user}---for example, using a
syntactic criterion or a termination prover.
If this criterion is not met, \cvc may answer \emph{sat} on
an unsatisfiable problem. For example, if we add the inconsistent definition
%
\begin{quote}
\small
\begin{verbatim}
(define-fun-rec e ((n Int)) Int (+ (e n) 1))
\end{verbatim}
\end{quote}
%
to the above problem, \cvc will still return ``satisfiable"
if \texttt{--fmf-fun} is specified.

>>>>

Admissibility is a semantic criterion that must be satisfied for each function
definition before applying the translation described in
Section~\ref{sec:encoding}. It is interesting it connect it with the more
standard notion of \emph{well-founded} function definitions, often called
``terminating'' definitions in a slight abuse of terminology. In such
definitions, all recursive calls are decreasing with respect to a well-founded
relation $r$, which must be supplied by the user or inferred automatically
using a termination prover. This ensures that the function is uniquely defined
at all points.

First-order logic has built-in notion of computation or termination. To ensure
that a function specification is well-founded, it is sufficient to require that
it would terminate when seen as a functional program for \relax{some}
evaluation order. For example,
$\{\forallf{\con{f}} n : {\typeint}.\;\, \lite\bigl(
n \leq 0,\allowbreak\; \con{p}( n ) \teq 1,\allowbreak\; \con{p}( n ) \teq 2 * \con{p}( n - 1 ) \bigr)\}$
can be shown well-founded under an evaluation strategy that consides one of the
two $\lite$ branches, based on the result of evaluating the condition.
Logically, such dependencies can be captured by congruence rules. Krauss
develops these ideas systematically in the more general context of higher-order
logic \cite[Section 2]{krauss-2009-phd}.

Another interesting fragment is the class of \emph{productive} corecursive
functions. Corecursive functions are functions to a coalgebraic datatype.
These functions can call themselves forever, without their being inconsistent.
Productive corecursive functions are functions that progressively reveal parts
of their (potentially infinite)
output \cite{turner-1995,abbott-et-al-2005,mcbride-productive}.
For example, given a type of infinite streams constructed by $\con{scons}$,
the definition
$\{\forallf{\const{e}} n.\;\, \const{e}(n) \teq \con{scons}(n,\; \const{e}(n + 1))\}$
falls within this fragment, since each call to $\const{e}$ produces one
constructor before entering the nested call. Like terminating recursion,
productive corecursion uniquely specifies a function. The following theorem
connects these notion with admissibility.

\begin{theorem}
Let 
\end{theorem}

@@

The \texttt{--fmf-fun} option instructs \cvc to assume that the
functions provided by the user are admissible.
Admissibility must be checked separately \emph{by the user}---for example, using a
syntactic criterion or a termination prover.
If this criterion is not met, \cvc may answer \emph{sat} on
an unsatisfiable problem. For example, if we add the inconsistent definition
%
\begin{quote}
\small
\begin{verbatim}
(define-fun-rec h ((n Int)) Int (+ (h n) 1))
\end{verbatim}
\end{quote}
%
to the above problem, \cvc will still return ``satisfiable"
if \texttt{--fmf-fun} is specified.
